---
title: "Ivy GAP -- Exploratory Adaptation Study"
author: "O. Celiku"
date: '`r format(Sys.time(), "%b %d %X %Y")`'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Intrinsic and exploratory runs inspection -- Examples

### Load intrinsic and exploratory walk examples 
```{r, message=FALSE, warning=FALSE}
library(PhenoExploreR)
data(IVY)

twopathways <- c("KEGG_B_CELL_RECEPTOR_SIGNALING_PATHWAY", "KEGG_DNA_REPLICATION")
twopathways.genes <- IVY@pathways[twopathways]
genes <- Reduce(union, twopathways.genes)
genes.indrange <- setdiff(match(genes, rownames(IVY@expr)),NA)
pathways.indrange <- match(twopathways, rownames(IVY@activity))
intrinsic <- readRDS("intrinsic_run_example.RDS")
# access walks
w.intrinsic <- intrinsic@w

exploratory <- readRDS("exploratory_run_example.RDS")
# access walks and goal
w.exploratory <- exploratory@w
w.goal <- exploratory@goal
```

### Look at genes and pathway walks of individual samples

#### First genes
These are plots at the single sample level (we select sample 4 below); for select genes look at individual genes walks (default indrange is the first 10 genes; set to desired indices). First intrinsic (each gene individually first, and then together):

```{r, message=FALSE, warning=FALSE, fig.width=8, fig.height=2.5}
sample_indx <- 4
w.intrinsic.ss <- w.intrinsic[[sample_indx]]
# genes walk matrix
w.intrinsic.ss.x <- w.intrinsic.ss@x
genesWalk(w.intrinsic.ss.x, scales="fixed")
```

```{r, message=FALSE, warning=FALSE, fig.width=4, fig.height=2.5}
allGenesWalk(w.intrinsic.ss.x, indrange=genes.indrange, ymin=-50, ymax=50)
```

Now exploratory:

```{r, message=FALSE, warning=FALSE, fig.width=8, fig.height=2.5}
w.exploratory.ss <- w.exploratory[[sample_indx]]
w.exploratory.ss.x <- w.exploratory.ss@x
genesWalk(w.exploratory.ss.x, scales="fixed")
```

```{r, message=FALSE, warning=FALSE, fig.width=4, fig.height=2.5}
allGenesWalk(w.exploratory.ss.x, indrange=genes.indrange, ymin=-50, ymax=50)
```

#### Now pathways 

Now repeat for pathways, at the single sample level, for select pathways: again set the pathway indices to inspect via indrange (default 1:10)

```{r, fig.width=8, fig.height=2}
# pathways walk matrix
w.intrinsic.ss.y <- w.intrinsic.ss@y
# use goal -- destination -- of exploratory to plot boundaries
pathwaysWalk(w.intrinsic.ss.y, bounds=w.goal, indrange=pathways.indrange, ncol=2, scales="fixed")

w.exploratory.ss.y <- w.exploratory.ss@y
pathwaysWalk(w.exploratory.ss.y, bounds=w.goal, indrange=pathways.indrange, ncol=2, scales="fixed")

```

```{r, message=FALSE, warning=FALSE, fig.width=4, fig.height=2.5}
allPathwaysWalk(w.exploratory.ss.y, indrange=pathways.indrange)
```

### Look at pathway distributions walks
```{r, message=FALSE, warning=FALSE, fig.width=8, fig.height=3}
distributionsWalk(w.intrinsic, startLoc="CT", destLoc = "CTmvp", IVY, pathways.indrange)
```
```{r, message=FALSE, warning=FALSE, fig.width=8, fig.height=3}
distributionsWalk(w.exploratory, startLoc="CT", destLoc = "CTmvp", IVY, pathways.indrange)
```

### Look at distances to the target distribution of the exploratory samples over time
```{r, warning=FALSE, message=FALSE, fig.width=6, fig.height=3}
ds <- lapply(w.exploratory, function(ww) ww@d)
allDistancesWalk(ds)
```

```{r}
sessionInfo()
```
