---
title: "Ivy GAP -- Exploratory Adaptation Study"
author: "O. Celiku"
date: '`r format(Sys.time(), "%b %d %X %Y")`'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Initial data exploration

## Load IVY data subset 
```{r, message=FALSE, warning=FALSE}
library(PhenoExploreR)
data(IVY)

require(GSVA)
require(reshape2)
require(made4)
require(factoextra)
require(FactoMineR)
require(knitr)
require(reshape2)
require(dplyr)
require(HistogramTools)
require(stringr)
require(ComplexHeatmap)
require(plot3D)
```

## Samples per location
```{r, message=FALSE, warning=FALSE}
table(IVY@design) %>% 
  as.data.frame %>%
  rename_all(funs(c("Location", "# Samples"))) %>% 
  kable
```

## Plot expression and pathway trajectories

First gene expression level, and then pathway activity level.
```{r, message=FALSE, warning=FALSE}
plotPhenotypePCA(IVY, method="BGA")
plotPhenotypePCA(IVY, expr=FALSE, method="BGA")
```
Standard PCA, first gene expression level, and then pathway activity level. 
```{r, message=FALSE, warning=FALSE}
plotPhenotypePCA(IVY, method="PCA")
plotPhenotypePCA(IVY, expr=FALSE, method="PCA")
```

## Pathway activity variance per location
```{r, message=FALSE, warning=FALSE, fig.height=2}
a <- IVY@activity
colnames(a) <- IVY@design

activity.variance <- a %>% melt %>% 
  transmute(pathway=Var1, location=Var2, activity=value) %>%
  group_by(location, pathway) %>% 
  summarise(sd=sd(activity))

activity.variance %>% 
  acast(location~pathway)%>% 
  apply(1, mean) %>% 
  round(2) %>%  
  as.data.frame %>% 
  rename_all(funs(c("Mean Variance"))) %>% kable

ggplot(activity.variance, aes(x=sd)) +
  geom_histogram(aes(y=..density..), colour="steelblue", fill="white", binwidth=0.01) +  
  geom_density(alpha=.2, fill="steelblue")  + theme_minimal() + facet_wrap(~location, ncol=5)

activity.variance %>% acast(location~pathway)%>%
Heatmap(row_title="Locations", column_title="Pathways", show_column_names = FALSE, show_row_names = TRUE, name="variance")
```


## Compute PDD and GDD for each location pair

First PDDs, and display distance distributions per location:
```{r, message=FALSE, warning=FALSE}
pdds <- computePDD(IVY@activity, design=IVY@design) 

pdds.melt <- pdds %>% melt %>% 
  as.data.frame() %>% 
  transmute(pathway=Var1, distance=value, comparison=Var2) %>% 
  mutate(location1=str_replace(comparison,"_.*",""), location2=str_replace(comparison, ".*_",""))

p <- ggplot(pdds.melt, aes(x=distance)) + 
  geom_histogram(aes(y=..density..), colour="steelblue", fill="white", binwidth=1) +   
  geom_density(alpha=.2, fill="steelblue")  + 
  theme_minimal() + 
  facet_wrap(~location1+location2, ncol=5)
p
```

Now GDDs:
```{r, message=FALSE, warning=FALSE}
gdds <- computePDD(IVY@activity, design=IVY@design, asGDD=TRUE) 

gdds %>% as.data.frame %>% 
  rename_all(funs(c("GDD"))) %>% 
  kable
```

GDD histogram: 
```{r, message=FALSE, warning=FALSE, echo=FALSE}
gdds.melt <- pdds.melt %>% as.data.frame %>% 
  group_by(location1,location2) %>% 
  summarise(GDD=sum(distance)) 
z <- matrix(NA, nrow=length(IVY@locations), ncol=length(IVY@locations))
colnames(z) <-  IVY@locations
rownames(z) <- IVY@locations
apply(gdds.melt, 1, function(x){
  z[x[1],x[2]] <<- x[3]
  z[x[2],x[1]] <<- x[3]
})
z[lower.tri(z)] <- NA

hist3D(z=apply(z, 2, as.numeric), theta=30, phi=30, axes=TRUE,label=TRUE, space=0.2,
       xlab="", ylab="", zlab="GDD", mar=c(5, 5, 5, 2), alpha=0.5, border ="gray", labels=rownames(z))
```

```{r}
sessionInfo()
```
