---
title: "Ivy GAP -- Exploratory Adaptation Study"
author: "O. Celiku"
date: '`r format(Sys.time(), "%b %d %X %Y")`'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Intrinsic and exploratory runs -- Examples

## Load IVY data subset 
```{r, message=FALSE, warning=FALSE}
library(PhenoExploreR)
data(IVY)

require(GSVA)
require(reshape2)
require(foreach)
require(doParallel)
require(assertthat)
```

# Examples of how to run intrinsic and exploratory runs

We recommend that these are run in batch form as the computations can require a long time, especially without utilizing parallel computation. Here we provide examples of how to run, and save results in .RDS files that can then be inspected separately.

## Example of intrinsic runs

Run an intrinsic run on a location of interest, for example CT; note that maxIter is set to 250, as the computation is time intensive. Also, we are using 8 cores and parallel computation. If this option is not available reduce numMaxCores to 1. outputDir specifies where individual samples walks can be saved, and is optional.

```{r, message=FALSE, warning=FALSE}
design <- IVY@design
design[design=="CT"] <- 0
design[design!=0] <- NA
phenoWandererRes <- wander(IVY@expr,design,IVY@motif,IVY@pathways,maxIter=250, outputDir="simulations",numMaxCores=8)
saveRDS(phenoWandererRes, "intrinsic_run_example.RDS")
```

## Example of exploratory runs

Choose 2 locations of interst, e.g. CT and CTmpv; run for 20 steps, with scaling paramater 0.95. 

```{r, message=FALSE, warning=FALSE}
design <- IVY@design
design[design=="CT"] <- 0
design[design=="CTmvp"] <- 1
design <- as.numeric(design)
phenoExplorerRes <- explore(IVY@expr,design,IVY@motif,IVY@pathways,maxIter=250, Jscaling=0.95, outputDir="simulations",numMaxCores=8)

saveRDS(phenoExplorerRes, "exploratory_run_example.RDS")
```

```{r}
sessionInfo()
```
